---
title: "Reproducible Research: Peer Graded Assignment 1"
output: html_document
---

##Mark Down for Peer Review
This an R markdown file generated for completion of the Coursera Reproducible Research Course Project 1. The following computations and visualizations are performed using the ggplot2, dplyr, and scales packages. https://www.coursera.org/learn/reproducible-research/.

##What is mean total number of steps taken per day?
The following code loads the file to a csv in the WD, converts date column to date datatype, and converts data set to a data frame.

```{r setup, include=FALSE}
activity <- read.csv("activity.csv")

activity$date <- as.Date(activity$date, "%Y-%m-%d")
activity <- as.data.frame(activity)
```

##Calculates mean and median of steps take by day

```{r}
steps <- with(activity, tapply(steps, date, sum, na.rm = TRUE))

mean(steps)
median(steps)
```

##What is the average daily activity pattern?

Using tapply, the means for each interval here are calculated across days.

```{r}
daymeans <- with(na.omit(activity), tapply(steps, interval, mean))
head(daymeans)
```

This code uses a logical vector to select the maximum value for the mean steps across intervals. The 835th interval gives the maximum value.

```{r}
daymeans[which(daymeans == max(daymeans))]
```

The following sums the na values and then reports the ratio of NA/total observations.

```{r}
library(scales)
sum(is.na(activity))
percent(sum(is.na(activity))/nrow(activity))
```

##Imputing missing values

My strategy for replacing NAs is to replace each NA value by steps mean per interval. This is accomplished by using a nested for loop to identify the respective interval of the row then replace the NA value with the mean for that interval.

The int and len variable are set up to manage the for loop sequences. The NAin and NA steps variables are sections of the data that will be used to replace the NA data after the loop.

```{r}
int <- unique(activity$interval)
len <- nrow(activity[is.na(activity),])

NAint <-  activity[is.na(activity),3]
NAsteps <- activity[is.na(activity),1]

for (j in 1:2304) {
       for (i in 1:288){
              if (NAint[j] == int[i])
                     NAsteps[j] <- daymeans[i]
       
       }
}

NAindex <- is.na(activity$steps)
activity$steps<- replace(activity$steps,NAindex, NAsteps)
```

As can be seen here, the NAs have been replaced by the relevant mean for the 5 minute interval.

```{r}
head(activity)
tail(activity)
```

steps take by day

```{r}
steps <- with(activity, tapply(steps, date, sum, na.rm = TRUE))

mean(steps)

median(steps)
```

Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

Yes they are changed, because the NAs were replaced based on the interval means, the mean and the median for the day now match up.

##Are there differences in activity patterns between weekdays and weekends?

First, this code creates a new factor variable.

```{r}
library(dplyr)
activity <- mutate(activity, day = weekdays(activity$date))

weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')

activity$day <- factor((weekdays(activity$date) %in% weekdays), 
         levels=c(FALSE, TRUE), labels=c('Weekend', 'Weekday'))
weekdays <- subset(activity, day == "Weekday")
weekends <- subset(activity, day == "Weekend")

weekendmeans <- with(weekends, tapply(steps, interval, mean))
weekdaymeans <- with(weekdays, tapply(steps, interval, mean))
head(weekendmeans)
head(weekdaymeans)
```







